---
title: "Problem Set 1"
subtitle: "INF 511"
author: "Muhammad"
fig-align: center
fig-cap-location: bottom
number-sections: true
format: 
    pdf: 
        documentclass: article
        geometry: 
          - top=1in
          - left=0.75in
          - bottom=1in
          - right=0.75in
editor: source
---

# $\chi^2$ disribution
## Write the probability distribution function.

A chi-square distributed random variable $X$ has $n$ degrees of freedom and follows a probability density function.
\begin{align*}
f(x) = \frac{x^{n/2-1}e^{-x/2}}{2^{n/2}\kern 0.08 em \Gamma(n/2)} \qquad \qquad x > 0,
\end{align*}
$for $n = 1, \, 2, \,  \ldots \$


```{r}
pdf <- dchisq(seq(0, 20, by=0.1),6)

```

## Plot the PDF.

```{r}
plot(seq(0, 20, by=0.1), pdf, type="l", xlab="x", ylab="f(x)")
```

# Discrete distribution 
For Question 2, the chosen distribution is the Binomial distribution.
\begin{align*}
\binom{n}{k} = \frac{\text{factorial}(n)}{\text{factorial}(k) \cdot \text{factorial}(n - k)}
\end{align*}


```{r}
n <- 10
p <- 0.5
x <- 0:n
k <- 6
binom_coef <- factorial(n) / (factorial(k) * factorial(n - k))
pmf <- binom_coef * p^k * (1 - p)^(n - k)
pmf
```

## Plot the PMF

```{r}
# Define the parameters of the binomial distribution
n <- 10
p <- 0.5

# Generate the values of x
x <- 0:n

# Calculate the PMF
pmf <- dbinom(x, n, p)

# Plot the PMF
plot(x, pmf, type = "l", xlab = "x", ylab = "P(x | n, p)")
points(x, pmf, pch = 19)

```



